Supplement to Automatic Matching of Bullet Lands

### 2016-01-06 imaging-supplement.Rnw
tons of cross sections to signatures

The scatterplot in Figure~\ref{fig:residual} shows the residuals of such a fit.
In this instance, the radius is estimated as $\hat{r} = \Sexpr{pars$radius}\mu m = \Sexpr{pars$radius/1000}mm$ and the land covers about \Sexpr{diff(range(dbr111$theta))} degrees.  Both of these estimates are consistent with a 9 mm bullet fired by a Ruger P-85.
The residuals are dominated, as expected, by the grooves, which show up as large positive residuals. For a surface cross cut at $x = 100$ there is a residual circular structure that does not show up for all cross sections.

@@

A single cylinder as a fit is unlikely to be a particularly good fit, because there seem to be quite massive deformations in the vertical direction. Even when we fit a circle to each cross section of the bullet, as in Figure~\ref{fig:circlefits}, this does not address all of these issues. While the wider circumference at the base of the bullet can be resolved by individual circular fits, the systematic residual structure in Figure~\ref{fig:residualb} stays the same.

@@

\begin{figure}[hbtp]
  \centering
<<circlefits, echo=FALSE, fig.width=10, fig.height=5, out.width='\\textwidth', warning=FALSE>>=
db2$land <- db2$part
db1$land <- db1$part
qplot(y, resid, data=subset(db2, x == 100), #colour=factor(x), 
      geom="line", size=I(.75), colour=I("grey70")) + 
  facet_wrap(~land, ncol=3, labeller="label_both") + 
  scale_colour_brewer(palette="Paired") +
  theme_bw() + theme(legend.position="bottom") + 
  geom_line(aes(y, resid, group = x), colour="black", size=.75, alpha=0.5,
             data = filter(db1, land==5, x == 100)[,c("y", "resid", "x")]) +
  ylab("Residuals from circular fit") +
  ggtitle("Bullet 1-5 in black")
@
\caption{Circular fit for each cross section of bullet 2, with bullet 1-5 overlaid.\label{fig:circlefits} Bullet 1-5 matches best with bullet 2-1.}
\end{figure}

@@

Figure~\ref{fig:crosscuts} shows a sequence of cross sections for bullet 1-5 (barrel 1) that are taken 50$\mu m$ apart, between 150$\mu m$  and 400$\mu m$. These are compared to a cross section at 100$\mu m$. Initially this comparison constitutes an almost perfect match between the two cross sections. However, the match quickly deteriorates with increasing distance to non-matches at a cross section of $x = 300\mu m$ or higher.  Only if the cross section is within 150$\mu m$ do we get a good visual match even when we know that the same bullet surface is being used. 
Given that we have to expect some variation in nominally the same cross section levels due to manual alignments in microscopes, we should take cross section levels into account in the automatic matching routine by evaluating matches at several cross sections. 
\begin{figure}[hbtp]
  \centering
<<crosscuts-vary, echo=FALSE, fig.width = 12, fig.height = 7, out.width = '\\textwidth', cache=TRUE, warning = FALSE>>=

@@

grid.arrange(lop[[1]], lop[[2]], lop[[3]], lop[[4]], lop[[5]], lop[[6]],
             ncol = 2)
@
\caption{\label{fig:crosscuts}Overview of the variations in the surface along different cross sections. The signature extracted at $x = 100$ is compared to signatures at every 50$\mu m$. With every step away from the original cross section site, the number of differences between the signatures increases; the number of maximum CMS decreases from initially 14 to  3 or fewer at $x = 300\mu m$ and higher. }
\end{figure}

>>>>

The scatterplot in Figure~\ref{fig:residual} shows the residuals of such a fit.
In this instance, the radius is estimated as $\hat{r} = \Sexpr{pars$radius}\mu m = \Sexpr{pars$radius/1000}mm$ and the land covers about \Sexpr{diff(range(dbr111$theta))} degrees.  Both of these estimates are consistent with a 9 mm bullet fired by a Ruger P-85.
The residuals are dominated, as expected, by the grooves, which show up as large positive residuals. For a surface cross cut at $x = 100$ there is a residual circular structure that does not show up for all signatures.

@@

A single cylinder as a fit is unlikely to be a particularly good fit, because there seem to be quite massive deformations in the vertical direction. Even when we fit a circle at each cross section of the bullet, as in Figure~\ref{fig:circlefits}, this does not address all of these issues. While the wider circumference at the base of the bullet can be resolved by individual circular fits, the systematic residual structure in Figure~\ref{fig:residualb} stays the same.

@@

\begin{figure}[hbtp]
  \centering
<<circlefits, echo=FALSE, fig.width=10, fig.height=5, out.width='\\textwidth', warning=FALSE>>=
db2$land <- db2$part
db1$land <- db1$part
qplot(y, resid, data=subset(db2, x == 100), #colour=factor(x), 
      geom="line", size=I(.75), colour=I("grey70")) + 
  facet_wrap(~land, ncol=3, labeller="label_both") + 
  scale_colour_brewer(palette="Paired") +
  theme_bw() + theme(legend.position="bottom") + 
  geom_line(aes(y, resid, group = x), colour="black", size=.75, alpha=0.5,
             data = filter(db1, land==5, x == 100)[,c("y", "resid", "x")]) +
  ylab("Residuals from circular fit") +
  ggtitle("Bullet 1-5 in black")
@
\caption{Circular fit to the signature of each land of bullet 2, with signature from bullet 1-5 overlaid.\label{fig:circlefits} The signature of bullet 1-5 matches best with bullet 2-1.}
\end{figure}

@@

Figure~\ref{fig:crosscuts} shows a sequence of signatures for bullet 1-5 (barrel 1) that are taken at cross sections 50$\mu m$ apart, between 150$\mu m$  and 400$\mu m$. These are compared to the signature at cross section 100$\mu m$. Initially, this comparison constitutes an almost perfect match between the two signatures. However, the match quickly deteriorates with increasing distance between the cross sections.  Only if signatures are from cross sections  within 150$\mu m$ do we get a good visual match even when we know that the same bullet surface is being used. 
Given that we have to expect some variation in nominally the same cross section levels due to manual alignments in microscopes, we should take cross section levels into account in the automatic matching routine by evaluating matches at several cross sections. 
\begin{figure}[hbtp]
  \centering
<<crosscuts-vary, echo=FALSE, fig.width = 12, fig.height = 7, out.width = '\\textwidth', cache=TRUE, warning = FALSE>>=

@@

grid.arrange(lop[[1]], lop[[2]], lop[[3]], lop[[4]], lop[[5]], lop[[6]],
             ncol = 2)
@
\caption{\label{fig:crosscuts}Overview of the variations in the surface along different cross sections. The signature extracted at $x = 100$ is compared to signatures at every 50$\mu m$. With every step away from the original cross section site, the number of differences between the signatures increases; the number of maximum CMS decreases from initially 22 to  four or fewer at $x = 300\mu m$ and higher. }
\end{figure}
